---
title: "Chiptole Locations"
author: "Sydney Yount"
date: "5/10/2022"
output: html_document
---

## Assignment Description

Create a web page using R Markdown that features a map created with Leaflet. 

Host your webpage on either GitHub Pages, RPubs, or NeoCities.

## Loading Required Packagers

```{r}
library(leaflet)
library(htmltools)
```
## Loading Data Set and Creating Data Frame

Goall: Map the location of every chiptole restaurant in the United States. Dataset includes the state, location, address, latitude and longitude. 

Source: https://www.kaggle.com/datasets/jeffreybraun/chipotle-locations?resource=download

```{r}
chipotleData <- read.csv(file = "chipotle_stores.csv",header=TRUE)
cDF <- data.frame(State = chipotleData$state, Location = chipotleData$location, Address = chipotleData$address, Latitude = chipotleData$latitude, Longitude = chipotleData$longitude)
```
## Creating Map using leaflet 
Ensure map has clusters due to high density of points on map.  
```{r}
map <- cDF %>%
  leaflet() %>%
  addTiles() %>%
  addMarkers(popup = paste("Location: ",htmlEscape(cDF$Location),"<br>Address: ",htmlEscape(cDF$Address)), clusterOptions = markerClusterOptions())
map
```